---
title: "R Markdown Presentation"
author: "Chuck Tucker"
date: "4/22/2020"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Creating New Slides

To create a new slide, just create a new heading.

## Some Basics

- Incorporate code inline
  - The result of 2 + 2 is `r 2+2`
  - Use 'r' as the first char inside the backticks

## Ordered Lists

1. Use numbers
2. If you want to create numbered lists that is

## Chunks of code

- Two ways to add code chunks:
1. Can click the *Insert* button to add a code chunk
2. Can do **Ctrl+Alt+i** on a PC or **Cmd+Shift+i** on Mac

## Displaying code in a chunk
- Use argument 'echo = TRUE' to display code + output
- Use argument 'echo = FALSE' to supress code and only display the output

```{r code_chunk_name, echo=FALSE}
print('This was printed by R')

```


## Including CSS

- Add argument `css: file.css` to your preamble to include a css file
- Easiest modications are using HTML tags


## RPubs

- Free way to publish R Markdown documents and presentations
- Publicly available
- Users can comment on the documents
- Click this link -> [RPubs](https://rpubs.com)
- To publish, click the publish button after knitting your document


## Shiny Apps

- This is a basic shiny interactive plot, built-in to the slide
- Shiny apps are not supported in static documents
  - Must add `runtime: shiny` to the preamble to include this functionality
  - The **Knit** button then becomes **Run Presentation**
- Shiny apps have a basic template

## My First Interactive Plot

- Interactive plot allowing you to set the exponent (x^input)

```{r polynomial_plot, echo=FALSE}
library(shiny)
shinyApp(
  ui= fluidPage(     # user interface is a fluidPage, allowing markdown/app to be displayed together
    sliderInput('exponent',      # name for the variable storing the slider value
                label = 'Exponent',     # label for this item on the UI
                min = 1,    # slider parameters including a starting 'value'
                max = 5,
                value = 2),
    plotOutput('polynomialPlot')   # assign name for this object to call in the server argument
  ),
  server = function(input, output) {    # create a function with input/output args
    output$polynomialPlot <- renderPlot(     # output must call renderPlot to be interactive
      curve(x^input$exponent, from = -5, to = 5)  # the plotting function using the slider variable
    )
  }
)
```




## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

